---
title: 'Geostatistical Data Example: Yearly Precipitation Anomaly Data'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>

\newcommand{\bm}{\mathbf}

# Dataset

```{r, echo=TRUE, fig.width=8, fig.align="center", message = FALSE, warning = F}
library(sp)
library(ggplot2)
library(maps)
library(mapdata)
library(geoR)
```

* Visualisation
  + 936 weather stations across Great Plains that are north of 40 degrees latitude
  + z: yearly precipitation anomaly  in 1962


```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
dfgeo = readRDS("datasets/dfgeo.rds")
states <- map_data("state")
plain <- subset(states, region %in% c("nebraska", "kansas", "colorado","wyoming", 
                                      "south dakota", "north dakota", "montana",
                                      "iowa", "minnesota", "missouri"))
gg2 = ggplot(data = plain) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.4)
gg2 + geom_point(data = dfgeo, aes(x = long, y = lat, color = z), size = 1)
```


* About the distance 
  + 1 degree in latitude = 111 km
  + 1 degree longitude at 45 degree latitude  $\approx$ 79 km
  + If you want a more accurate distance than the above approximation, you need to use the function *rdist.earth*. 

```{r, echo=TRUE, fig.width=8, fig.align="center"}
dfgeo$x = dfgeo$long * 79
dfgeo$y = dfgeo$lat * 111
```

The dataset is divided into two part, where *dfgeo_train* is used for modelling, and *dfgeo_test* is for kriging. 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
idx = seq(1, length(dfgeo$x), 5)
dfgeo_test = dfgeo[idx, ]
dfgeo_train = dfgeo[-idx,]
```


# Model 1

* For Model 1: $$
Z(\mathbf{s}_i)=\mu+U(\bm{s}_i)+\nu(\bm{s}_i)
%\sum_{j=1}^p \beta_j d_j(x)+S(x)+Z(x) $$
  + Constant mean: $\mu$
  + Covariance function: exponential 

## Variogram and Initial Value for MLE

* The function *variog* is used to calculate the semivariogram in our slides. 
  + option: "bin"
  + max.dist: maximum distance for the semivariogram
  


```{r, echo=TRUE, fig.width=8, fig.align="center"}
dfgeodata =  as.geodata(dfgeo_train, coords.col=5:6, data.col=4, covar.col = 1:3)
plot(variog(dfgeodata, option="bin", max.dist=900), 
     xlab = "h", ylab = "variogram")
```

For the initial value, you first need to decide what type of covariance function you want to use. Here, we use the exponential covariance function (i.e., the matern class with $\kappa = 0.5$.)

For the exponential covariance function, we choose the initial value $(\sigma^2, \phi, \tau^2) = (0.8, 200, 0.1)$. Since the line roughly follow the circles (the semivariograms), $(\sigma^2, \phi, \tau^2) = (0.8, 200, 0.1)$ is an appropriate  initial value. 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
plot(variog(dfgeodata, option="bin", max.dist=900), 
     xlab = "h", ylab = "variogram")
lines.variomodel(seq(0, 900, l = 100), 
                 cov.pars = c(0.8, 200),           
                 cov.model = "mat", kap = 0.5, nug = 0.1)
```

## Model Specification

* The likfit function is used for MLE and REML for the geostatistical model. To understand how the covariance function is specified, you first need to know the *default* setting. 
  + cov.model= "matern": matern class
  + kappa = 0.5
  + fix.kappa = TRUE: the kappa parameter is fixed at kappa = 0.5
  
* Therefore, if you do not write anything in the function, the default covariance function is the exponential covariance function, as shown in Model 1. 

\midskip

* Next, whether to estimate the nugget effect $\tau^2$. The default setting is 
  + fix.nugget = FALSE: the nugget effect is estimated. If you want to fix the nugget effect, you need to write *fix.nugget = TRUE*. 
* Therefore, 
  + nug = 0.1: the nugget effect is estimated, and the initial value is 0.1, as shown in Model 1. 
  + fix.nugget = FALSE, nug = 0.1: the same as the above.
  + fix.nugget = TRUE, nug = 0.1: the nugget effect is fixed (NOT estimated) and the value is fixed at 0.1. 

  
## Model Fitting



```{r, echo=TRUE, fig.width=8, fig.align="center"}
## MLE: constant trend + exponential covariance function. 
m1 = likfit(dfgeodata, ini = c(0.8, 200), nug = 0.1)
```

First, check the convergence message, that is, $convergence = 0 indicates a successful convergence.  (The likfit uses the *optim* funciton to minimize, that is, the message is the same as the message from the *optim* function)
```{r, echo=TRUE, fig.width=8, fig.align="center"}
m1$info.minimisation.function
```

Second, whether the model improves over the linear model. At the bottom of the summary table, there are two rows: *Maximised Likelihood* (for the spatial model) and *non spatial model*.  



## The Results

Summary table includes most of information. 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
summary(m1)
```

* The $\mu$ part
  + in the summary table, it is denoted as *\beta*. In our example, $\hat \mu = -0.2176$. 
  + It can also obtained by *m1$beta*
  + The estimated asymptotic variance (matrix) is *m1$beta.var*
  + A confidence interval or hypothesis testing can be conducted. 
  + For example, $H_0 = 1$ versus $H_1 \neq 1$. The test statistic = *(m1\$beta - 1)/sqrt(m1\$beta.var)* (approximately)$\sim$ N(0,1). 
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
m1$beta
# the estimated covariance matrix of beta. 
m1$beta.var
# test statistic for the example
(m1$beta - 1)/sqrt(m1$beta.var)
```

* The $\bm\theta$ part
  + The estimate can be found in the summary table or via the codes below. 
  + The estimated asymptotic covariance matrix can be calculated using the formula in our slides, but it is not available through the *likfit* function.
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
# cov.pars
m1$cov.pars
# partial sill 
m1$sigmasq
# the range parameter
m1$phi
# nugget
m1$nugget
```
 

* Other information
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
# AIC of the model
AIC(m1)
# and others...
m1$practicalRange
logLik(m1)
```

# Other Models

* For Model 2: $$
Z(\mathbf{s}_i)=\mu+U(\bm{s}_i)+\nu(\bm{s}_i)
%\sum_{j=1}^p \beta_j d_j(x)+S(x)+Z(x) $$
  + Constant mean: $\mu$
  + Covariance function: matern class with kappa = 1.5 


Some key codes
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
## MLE: constant trend + matern class with kappa = 1.5
#plot(variog(dfgeodata, option="bin", max.dist=900), 
#     xlab = "h", ylab = "variogram")
#lines.variomodel(seq(0, 900, l = 100), 
#                 cov.pars = c(0.8, 150),           
#                 cov.model = "mat", kap = 1.5, nug = 0.1)


#m2 = likfit(dfgeodata, ini = c(0.8, 150), nug = 0.1, cov.model = "mat", kap = 1.5)
#m2$info.minimisation.function$convergence
#summary(m2)
```

* For Model 3: $$
Z(\mathbf{s}_i)=\beta_0+x_1(\mathbf{s}_i)\beta_1+\nu(\bm{s}_i)
%\sum_{j=1}^p \beta_j d_j(x)+S(x)+Z(x) $$
  + Linear trend: $x_1$ is the elevation
  + Covariance function: exponential

The model is fitted using the REML method. 

Some key codes
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
## The variogram, note ``trend = ~elev"
#plot(variog(dfgeodata, option="bin", max.dist=900, trend = ~elev), 
#     xlab = "h", ylab = "variogram")
#lines.variomodel(seq(0, 900, l = 100), 
#                 cov.pars = c(0.8, 200),           
#                 cov.model = "mat", kap = 0.5, nug = 0.1)

#m3 = likfit(dfgeodata, ini = c(0.8, 200), nug = 0.1, trend = ~ elev, lik.method = "REML")
```



# Kriging

There are two key functions for the kriging
  + krige.control: specify the parameter for kriging, no computing involved.
  + krige.conv: to compute 

\vspace{5pt}

* Usage
  + krige.control(type.krige = "ok", obj.model = NULL)
  + Type of krige: "sk", "ok" (default)
    + sk: simple kriging
    + ok: ordinary kriging and universal kriging
  + obj.model: typically an output of *likefit*

\vspace{5pt}

* Usage
  + krige.conv(geodata, locations, krige)
  + geodata: the dataset for the model fitting
  + locations: locations for predicting $\mathbf s_0$, data.frame (can be multiple locations.) 
  + krige: output of krige.control.


## Simple Kriging
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
kc1 = krige.control(obj.model = m1, type.krige = "SK")
m1krig_sk = krige.conv(dfgeodata, krige = kc1, locations = dfgeo_test[,5:6])
```

## Ordinary Kriging
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
kc2 = krige.control(obj.model = m1)
m1krig_ok = krige.conv(dfgeodata, krige = kc2, locations = dfgeo_test[,5:6])
```

```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
# The kriging predictors are the same
rbind(m1krig_sk$predict[1:5], m1krig_ok$predict[1:5])
# The kriging variances are different
rbind(m1krig_sk$krige.var[1:5], m1krig_ok$krige.var[1:5])
```


## Universal Kriging
* We need to provide $x_1(\mathbf s_0), \ldots, x_p(\mathbf s_0)$ for the function
  + in function *krige.control*
  + trend.d: specify data trend
  + trend.l: specify prediction trend

We take Model 3 as an example  
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
#kc3 = krige.control(type.krige = "ok", obj.model = m3, trend.d = ~elev, trend.l = ~dfgeo_test[,3])
#m3pred = krige.conv(dfgeodata, krige = kc3, locations = dfgeo_test[,5:6])
```
