---
title: 'Spatial Lag Data Example: Used Car Data'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>

\newcommand{\bm}{\mathbf}

# Dataset

```{r, echo=TRUE, fig.width=8, fig.align="center", message = FALSE, warning = F}
library(ggplot2)
library(sf)
library(maps)
library(dplyr)
library(spData)
library(spdep)
library(spatialreg)
```


* We want to study the relationship between price of used cars and the taxes charged across 48 states in 1960
  + tax.charges: taxes and delivery charges for 1955-9 new cars
  + price.1960: 1960 used car prices by state


```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
data(used.cars)
head(used.cars)
```





# Visualisation

First, we read into the us states maps and convert it the *sf* object.
```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
state_sf = sf::st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))
```


Note that the state names in *used_cars* are abbreviation while full names in *state\_sf*. 
We want to combine these datasets together by names, and used *state.csv* which includes both abbreviation and full names of all states. 

```{r, echo=TRUE, fig.width=8, fig.align="center"}
abb = read.csv("datasets/states.csv")
abb = abb |> mutate(state_lower = tolower(State))
used.cars$Abbreviation = rownames(used.cars)

head(abb)
head(state_sf)
```

```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
used = state_sf |> rename(state_lower = ID) |> left_join(abb, by = "state_lower") |> right_join(used.cars, by = "Abbreviation") |> rename(price = price.1960, tax = tax.charges)
```

Next, we visualise the price variable and tax variable. We can see there is a spatial dependence for both variables. 

```{r, echo=TRUE, fig.width=8, fig.align="center"}
ggplot() + geom_sf(data = used, aes(fill = price))
ggplot() + geom_sf(data = used, aes(fill = tax))
```

## Spatial Dependence for price

```{r, echo=TRUE, fig.width=8, fig.align="center"}
moran.test(used$price, nb2listw(usa48.nb))
moran.plot(used$price, nb2listw(usa48.nb),
           labels=used$Abbreviation)
```

# Modelling 

## OLS

If we fit a linear model, we can see the resulting residuals are still spatially correlaed. 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
mod_ls <- lm(price ~ tax, data=used)
summary(mod_ls)

used$res = mod_ls$residuals
ggplot() + geom_sf(data = used, aes(fill = res))
```

## Spatial Lag Model




```{r, echo=TRUE, fig.width=8, fig.align="center"}
mod_lag <- lagsarlm(price ~ tax, data=used,
                   nb2listw(usa48.nb))
summary(mod_lag)
```




```{r, echo=TRUE, fig.width=8, fig.align="center"}
used$lagres = mod_lag$residuals
ggplot() + geom_sf(data = used, aes(fill = lagres))
```