---
title: 'Geom_polygon -- Maps'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>





# USA map
R package *maps* and *mapdata* provide some maps. We will use them for the USA map
```{r, echo=TRUE, fig.width=6, fig.align="center", warning = F, message = F}
library(ggplot2)
library(maps)
library(mapdata)
library(sp)
usa <- map_data("usa")
dim(usa)
head(usa)
tail(usa)
```

First, we plot usa map

```{r, echo=TRUE, fig.width=10, fig.align="center"}
library(gridExtra)
gg0 = ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
  coord_quickmap()
gg1 <- ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = "violet", color = "blue") + 
  coord_quickmap()
labs <- data.frame(
  long = c(-122.064873, -122.306417),
  lat = c(36.951968, 47.644855),
  names = c("SWFSC-FED", "NWFSC"),
  stringsAsFactors = FALSE
)  
gg2 = gg1 + 
  geom_point(data = labs, aes(x = long, y = lat), color = "yellow", size = 2)

grid.arrange(gg0, gg1, gg2, nrow = 1)
```

* Key points
  + "group = group" is essential, you can try code without the group argument. 
  + coord_quickmap(): one degree longitude is not the same distance as one degree latitude. For the latitude at 40, it is about $1/cos(2\pi/9) \approx 1.3$ of that of the longitude. coord_quickmap() can fix this automatically, see Chpater 9.7 of RDS book (https://r4ds.hadley.nz/layers#coordinate-systems). 
  + the two dots represents two cities. 






# US state
```{r, echo=TRUE, fig.width=10, fig.align="center", warning = F}
### Draw plots for all USA States
states <- map_data("state")
ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_quickmap() + guides(fill=FALSE) 

### Draw a subset of USA states
### Use function subset to select states. 
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))
ggplot(data = west_coast) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = "palegreen", color = "red") + 
  coord_quickmap()
```

Can you tell the difference between *color* and *fill* argument? 



