---
title: "Solutions"
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>


# Week 3

## Visualize the daily pattern of *mel_sel* data. 

```{r, echo=TRUE, fig.width=8, fig.align="center", message = FALSE}
library(lubridate)
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r, echo=TRUE, fig.width=8, fig.align="center", message = FALSE}
load("datasets/melb_sel.Rdata")
head(melb_sel)

mel_d_avg = melb_sel |> mutate(wday = wday(date, label = TRUE), hour = hour(date)) |> group_by(hour) |>
  summarise(avg_counts = mean(counts)) 
  
ggplot(mel_d_avg, aes(x = hour, y = avg_counts)) + geom_line()
```

## (Optional) If you want to visulize the weekly pattern. 

```{r, echo=TRUE, fig.width=8, fig.align="center", message = FALSE}
melb_w_avg = melb_sel |> mutate(wday = wday(date, label = TRUE), hour = hour(date)) |> 
  group_by(wday, hour) |> summarise(avg_counts = mean(counts))
melb_w_avg$mdayh = melb_sel$date[1:168]
melb_w_avg |> 
  ggplot(aes(x = mdayh, y = avg_counts)) + geom_line() + 
  scale_x_datetime(date_labels = "%a %H:%M:%S", date_breaks = "1 day")
```

# Week 4


Question: What is the expected number of events?

Answer: 
\[\lambda \nu(D) = 100\times 4 = 400. \]



Question: What do you conclude from the above plot?

Answer: We fail to reject the null hypothesis since the estimated G function is inside the envelope. 


Question: Can you test *rNS* data whether it follow clustered point pattern?

```{r echo=TRUE, results="hold"}
library(spatstat)
set.seed(2025)
nclust <- function(x0, y0, radius, n) {
          return(runifdisc(n, radius, centre=c(x0, y0)))
            }
rNS = rNeymanScott(kappa=10, expand=0.1, 
                   rcluster = nclust, radius=0.1, n=10)
r=seq(0,1, by=0.005)
Gtest=envelope(rNS,fun=Gest, r=r, nrank=5, nsim=200)
plot(Gtest, xlim = c(0,0.1))
```


```{r echo=TRUE, results="hold"}
quadrat.test(rNS, nx = 4, ny= 4, alternative = "clustered", method = "Chisq")
```

Both methods conclude *rNS* data follow a clustered point pattern. 


# Week 8

```{r, echo=TRUE, fig.width=8, fig.align="center"}
library(geoR)
dfgeo = readRDS("datasets/dfgeo.rds")
dfgeo$x = dfgeo$long * 79
dfgeo$y = dfgeo$lat * 111

idx = seq(1, length(dfgeo$x), 5)
dfgeo_test = dfgeo[idx, ]
dfgeo_train = dfgeo[-idx,]

dfgeodata =  as.geodata(dfgeo_train, coords.col=5:6, data.col=4, covar.col = 1:3)


plot(variog(dfgeodata, option="bin", max.dist=900), 
     xlab = "h", ylab = "variogram")
lines.variomodel(seq(0, 900, l = 100), 
                 cov.pars = c(0.8, 200),           
                 cov.model = "mat", kap = 0.6, nug = 0.1)
```

```{r, echo=TRUE, fig.width=8, fig.align="center"}
m4 = likfit(dfgeodata, ini = c(0.8, 200), nug = 0.1, cov.model = "matern", fix.kappa = FALSE, kappa = 0.6)
summary(m4)
```