---
title: 'Week 7: Geostatistics--Estimation'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>

<span style="color: red;">Due to the Easter Break, this Lab will be held in Week 9. </span>


# Introduction

* In this Lab, we will learn
  + Semivariogram: Cp 4 -- "Introduction and Semivariogram"
  + MLE: Cp 4 -- "Covariance and Semivariogram Models", "Paramatric Modelling: MLE"; Cp 5 -- "MLE"



# Soil Data & Semivariogram


## Soil Example

$Z(\mathbf{s})$: calcium content in soil at $\mathbf{s}$. 
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(geoR)
data(ca20)
points(ca20,cex.min = 0.2, cex.max = 1.5, pch = 21)
```




## Semivariogram
```{r echo=TRUE, message=FALSE, warning=FALSE, hold=TRUE}
plot(variog(ca20, option="bin", max.dist=510), 
     xlab = "h", ylab = "gamma(h)")
```

Here, the function *variog* is used to calculate the semivariogram in our slides. 


* Function: variog
  + argument
  + option: "bin"
  + max.dist: maximum distance for the semivariogram
  


# MLE


## MLE: Initial Value

<span style="color: red;">"Eye-fit"</span>

```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
plot(variog(ca20, option="bin", max.dist=510), 
     xlab = "h", ylab = "gamma(h)")
lines.variomodel(seq(0, 510, l = 100), 
                 cov.pars = c(100, 120),           
                 cov.model = "mat", kap = 0.5, nug = 30)
```

Based on the plot, the initial value $(\sigma^2, \phi, \tau^2) = (100, 120, 30)$.


## MLE: Function
```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
# trend - constant mean
m1 = likfit(ca20, ini = c(100, 120), nug = 30)
```

```{r echo=TRUE, message=FALSE, warning=FALSE, hold=TRUE}
# trend - constant mean
summary(m1)
```

* Results of output
  + Spatial-dependence parameter estiamtes: $(\sigma^2, \phi, \tau^2)$ 
  + no standard error is provided.
  + cov.pars: $(\sigma^2, \phi)$ 
  + nugget: $\tau^2$
  + kappa


The output of MLE
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
# cov.pars
m1$cov.pars
# partial sill 
m1$sigmasq
# the range parameter
m1$phi
# nugget
m1$nugget
```


* <span style="color: red;">The arguments of function</span> *likfit*
  + argument
  + cov.model: "exp" (default), more options see cov.pars
  + Fixed or Estimated: fix.nugget = FALSE, fix.kappa = TRUE, kappa =0.5
  + Initial value: ini.cov.pars, nugget 

\vspace{5pt}

* Remarks: we have 4 parameters: nugget $\tau^2$, partial sill $\sigma^2$, range parameter $\phi$, smooth parameter $\kappa$ ( $\nu$ in our slide. )
  + Function allow us either to pre-specify (fix) or estimate it
  + Most of the time, $\kappa$ is fixed.
  
  
Here are some exmaple: 

* Example 1: no specification.  "likfit(ca20, ini = c(100, 200), nug = 50)"
  + Exp model ($\kappa$ = 0.5 fixed). 
  + To estimate: $\tau^2$, $\sigma^2$, $\phi$
  
\vspace{5pt}

* Example 2: no specification.  "likfit(ca20, ini = c(100, 200), nug = 50, fix.nugget=TRUE)"
  + Fixed: $\kappa$ = 0.5, nugget $\tau^2 = 50$
  + To estiamte: $\sigma^2$, $\phi$


Many other informatons can be obtained. 
```{r echo=TRUE, message=FALSE, warning=FALSE, results="hold"}
# the estimated covariance matrix of beta. 
m1$beta.var
# AIC of the model
AIC(m1)
# and others...
m1$practicalRange
logLik(m1)
```
You can also study possible covariance functions by reading details (accessed by typing "?cov.spatial")
  
## Spatial Trend

### Two ways
  + Specify directly in function like "likfit": `fit = likfit(ca20, ini = c(100, 60), trend = ~altitude)`
  + Specify it through function "trend.spatial": See below

\vspace{5pt}

* Usage: only to specify, not computing involved in this function.
  + `trend.spatial(trend, geodata)`

\vspace{5pt}

* Argument *trend*
  + trend = "cte": constant mean
  + trend = ~altitude: $\mu(s)=\beta_0+\beta_1 altitude$
  + Other options: "1st", "2nd", see help files for more details





# Geodata format

The package *geoR* requires *geodata* object, and in this part, we show how to create geodata object

* geodata object: a list with two obligatory components
  + coords: a matrix with the two-dimensional coordinates of the sampling locations
  + data: a vector (vectors) giving the corresponding measured value(s) at each of sampling location. 
  

## Create geodata

* <span style="color: red;">Converting a data.frame to geodata</span>: many functions in geoR requires geodata format.
  + *as.geodata(obj, coords.col = 1:2, data.col = 3, covar.col = NULL, covar.names = "obj.names")*
  + coords.col = 1:2 (default): the first two columns are x-axis and y-axis. 
  + data.col = 3 (default): the column for the response ($Z(\mathbf{s})$). The default is the third column, but you can specify it. In the following example, the 5th column is for the response. 
  + covar.col: columns for covariates. 
```{r message=TRUE, echo=TRUE, warning=FALSE}
library(geoR)
data(camg)     # data.frame format
# geodata format
geocamg = as.geodata(camg, data.col=5, covar.col=c(3,6), 
                     covar.names = c("elevation", "mg20"))    
```
