---
title: 'Week 6: Visualisation of Spatial Data -- Maps'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>

<span style="color: red;">Due to the Easter Break, this Lab will be held in Week 7. </span>

# Introduction
* In this Lab, we will learn to draw some maps. 





# USA map
R package *maps* and *mapdata* provide some maps. We will use them for the USA map
```{r, echo=TRUE, fig.width=6, fig.align="center", warning = F, message = F}
library(ggplot2)
library(sf)
library(maps)
library(mapdata)
library(sp)
usa <- map_data("usa")
dim(usa)
head(usa)
tail(usa)
```

First, we plot usa map
```{r, echo=TRUE, fig.width=10, fig.align="center"}
library(gridExtra)
gg0 = ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3)
gg1 <- ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = "violet", color = "blue") + 
  coord_fixed(1.3)
labs <- data.frame(
  long = c(-122.064873, -122.306417),
  lat = c(36.951968, 47.644855),
  names = c("SWFSC-FED", "NWFSC"),
  stringsAsFactors = FALSE
)  
gg2 = gg1 + 
  geom_point(data = labs, aes(x = long, y = lat), color = "yellow", size = 2)

grid.arrange(gg0, gg1, gg2, nrow = 1)
```

* Key points
  + "group = group" is essential, you can try code without the group argument
  + coord_fixed(): It fixes the relationship between one unit in the y direction and one unit in the x direction. 



## US state
```{r, echo=TRUE, fig.width=10, fig.align="center", warning = F}
### Draw plots for all USA States
states <- map_data("state")
ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_fixed(1.3) + guides(fill=FALSE) 

### Draw a subset of USA states
### Use function subset to select states. 
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))
ggplot(data = west_coast) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = "palegreen", color = "red") + 
  coord_fixed(1.3)
```

Can you tell the difference between *color* and *fill* argument? 





# Victoria maps

* Most shapefile can found in https://data.gov.au. 
+ All LGA: VICLGA folder -- 92 elements
+ All suburb: VICSub folder  -- 2973 elements

All datasets are read into as *sf* objects
```{r, echo=TRUE, fig.width=6, fig.align="center", message = F}
rm(list = ls())
library(sf)
VICLGA = st_read(dsn = "datasets/VICLGA/VIC_LOCALITY_POLYGON_shp.shp", "VIC_LOCALITY_POLYGON_shp")
VICSub = st_read(dsn="datasets/VICSub/VIC_LOCALITY_POLYGON_shp.shp", "VIC_LOCALITY_POLYGON_shp")
```



First, we make sure that all *sf* objects has the same CRS
```{r, echo=TRUE, fig.width=6, fig.align="center", message = F}
## Retrieve CRS
st_crs(VICLGA) 
st_crs(VICSub)

## Transform crs system to WGS84(epsg: 4326)
VICLGA <- st_transform(VICLGA, 4326)
VICSub <- st_transform(VICSub, 4326)

## Check CRS
#st_crs(VICLGA) 
#st_crs(VICSub)
```

Second, we want to obtain all suburbs in City of Banyule. Note that we already have all suburbs in Victoria, and we also know the boundary of City of Banyule from VICLGA. We only need to ``cut" all suburbs from *VICSub* using boundary of City of Banyule. 

* One useful funciton in *st_intersection* in package *sf*
  + st_intersection(x, y)
  + this function returns a geographic subset of an object *x* as specified by an Extent object *y*

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE, warning = F}
library(raster)
library(mapview)


## Obtain all suburbs in City of Banyule (including part of suburb)
i = which(VICLGA$VIC_LGA__3 == "BANYULE")
CMEL = VICLGA[i, ]
tmp = st_intersection(VICSub, CMEL)
tmp$VIC_LOCA_2
```


The object *tmp* includes suburbs which is not in City of Banyule, such as Balwyn North. Here is a list of all suburbs in City of Banyule (https://en.wikipedia.org/wiki/City_of_Banyule). The reason for this is likely due to the different resolution and accuracy of shapefiles (VICSub and VICLGA).   

Therefore, I will remove all the suburbs with a very small area (Here, the unit is square meter, and 10000 is small by a suburb standard.)  

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
## Obtain all suburbs in City of Banyule (including part of suburb)
tmp$area = st_area(tmp)
BSub = tmp[as.numeric(tmp$area) > 100000,]
mapView(BSub)
save(BSub, file = "datasets/BSub.Rdata")   
```

There are more shapefiles in *PlanMel* Folder, which outlines some future plan in greater Melbourne regions and can be downloaded from https://www.planmelbourne.vic.gov.au/maps/spatial-data.

There are major activity centres. According to the goverment, *These are places that provide a suburban focal point for services, employment, housing, public transport and social interaction. They have different attributes and provide different functions, with some serving larger subregional catchments. Plan Melbourne identifies 121 existing and future Major Activity Centres across Melbourne.*

Suppose we want to obtain all major activity centres in the City of Banyule. Note these acitivity centres are represented as *points*

```{r, echo=TRUE, fig.width=7, fig.align="center", message = FALSE}
centres = st_read(dsn = "datasets/PlanMel/Activity Centres/ActivityCentrePoints_font_point.shp", 
                 "ActivityCentrePoints_font_point")
centres <- st_transform(centres, 4326)
## obtain all centres within the City of Banyule (CMEL)
tmp = st_within(centres, CMEL, sparse = F)
Bcen = centres[tmp[,1],]

library(ggplot2)
library(sf)
ggplot() + geom_sf(data = BSub, aes(fill = VIC_LOCA_2)) + 
  geom_sf_label(data = BSub, aes(label = VIC_LOCA_2), size = 2.1) + 
  geom_sf(data = Bcen, size = 4) + theme(legend.position="none")
```




Suppose you also want to draw major roads in City of Banyule. The shapefile of all roads are in the folder *VMTRANS*. Here, we want to draw all roads with Road Class = 0, 1, 2. (I do not know what Road Class really means here, but it appears the smaller number indicates a more important road.)


```{r, echo=TRUE, fig.width=7, fig.align="center", message = FALSE}
road = st_read(dsn = "datasets/VMTRANS/TR_ROAD.shp", "TR_ROAD")
road <- st_transform(road, 4326)
## You can see all roads in Banyule
#mapView(road)
## We only draw road with CLASS_CODE = 0, 1, 2
highway = road[road$CLASS_CODE < 3,]
#mapView(highway)

## Now add all together
ggplot() + geom_sf(data = BSub, color = "gray") + 
  geom_sf(data = Bcen, size = 4) + 
  geom_sf(data = highway, aes(color = as.factor(CLASS_CODE)), lwd = 1.5)  
```



