---
title: 'Week 2: Visualisation of Spatial Data--ggplot2'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>


```
The following materials are partly based on "https://eriqande.github.io/rep-res-eeb-2017/" 
by Eric C. Anderson, Kristen C. Ruegg, Tina Cheng.
```

# Introduction
* In this Lab, we will learn the following functions in ggplot2. 
  + ggplot
  + geom_point: draw points
  + geom_line: draw lines
  + geom_polygon: draw polygons
  + geom_sf: draw sf class


# ggplot2
According to wikipedia and R-project website, 
*``ggplot2 is a data visualization package for the statistical programming language R. Created by Hadley Wickham in 2005, ggplot2 is an implementation of Leland Wilkinson's Gurky Gang—a general scheme for data visualization which breaks up graphs into semantic components such as scales and layers. ggplot2 can serve as a replacement for the base graphics in R and contains a number of defaults for web and print display of common scales. Since 2005, ggplot2 has grown in use to become one of the most popular R packages."*

It is a general framework for data visualisation, and in this class, we will use it to visualise spatial data and spatio-temporal data. 




## pv dataset

The pv dataset includes information of  the history of pole-vaulting world record "http://en.wikipedia.org/wiki/Men's_pole_vault_world_record_progression". Here, the data is already stored in *pv.RData*. 

First, read the RData set. 
```{r, echo=TRUE, fig.width=6, fig.align="center", warning = F}

library(ggplot2)
load("datasets/pv.RData")
```

Here are some plots. 
```{r, echo=TRUE, fig.width=3, fig.height =3, fig.align="center"}
par(mfrow=c(1,3))      # This command will not work in ggplot2
## Tell ggplot the data, x variable and y variable
g <- ggplot(data = pv, aes(x = Date, y = Meters))
g    
## Add points
g2 = g + geom_point()
g2
## Add red lines
g3 <- g2 + geom_line(color = "red")
g3

## Change the color of a point by adding a more layer
g4 <- g3 + geom_point(data = pv[4,], color="red", fill = "blue", shape = 21, size = 4, stroke =2) 
g4  
```




* The best way to learn trying different codes and see what it plots. Here are some suggestions
  + ggplot()
  + ggplot(data = pv, aes(x = Date, y = Meters))
  + ggplot()+geom_point(data = pv, mapping = aes(x = Date, y = Meters),col=2)
  + For more details on specification, type *vignette("ggplot2-specs")*

* ggplot:  initialize the plotting and everything set up in *aes* is valid for all subsequent plots. 
  + Try g <- ggplot(data = pv, aes(x = Date, y = Meters, col = 3)), and repeat everything above. 
  + More details can be found https://www.datanovia.com/en/blog/ggplot-aes-how-to-assign-aesthetics-in-ggplot2/ 
  
* Add layers. 
  + geom_point: more example at https://ggplot2.tidyverse.org/reference/geom_point.html
  + geom_line: https://ggplot2.tidyverse.org/reference/geom_path.html?q=geom%20_%20lines#arguments

## More examples

This part is to provide some interesting plots
```{r, echo=TRUE, fig.width=9, fig.height=3, fig.align="center"}
library(gridExtra)    # for the function grid.arrange
d <- ggplot(data = pv, aes(x = Date, y = Meters))
d2 <- d + geom_ribbon(aes(ymax = Meters), ymin = min(pv$Meters), alpha = 0.4, fill = "turquoise")
d3 <- d2 + geom_line(color = "blue")
d4 <- d3 + geom_point(color = "orange")
grid.arrange(d2, d3, d4, nrow = 1)
d5 <- d4 + geom_rug(sides = "bl", mapping = aes(color = Nation))
d5
```

For, *geom_ribbon*, try to see
```{r, echo=TRUE, fig.width=9, fig.height=3, fig.align="center"}
dr1 = d + geom_ribbon(aes(ymax = Meters), ymin = min(pv$Meters), fill = "grey70") + geom_line(aes(y = Meters))
dr2 = d + geom_ribbon(aes(ymin = Meters - 0.5, ymax = Meters + 0.5), fill = "grey70") + geom_line(aes(y = Meters))+ geom_point()
dr3 = d + geom_ribbon(aes(ymin = Meters), ymax = max(pv$Meters), fill = "grey70") + geom_line(aes(y = Meters))
grid.arrange(dr1, dr2, dr3, nrow = 1)
```



# ggplot for spatial data

R package *maps* and *mapdata* provide some maps. 

## USA map
We will use the USA map
```{r, echo=TRUE, fig.width=6, fig.align="center"}
library(maps)
library(mapdata)
library(sp)
usa <- map_data("usa")
dim(usa)
head(usa)
tail(usa)
```

First, we plot usa map
```{r, echo=TRUE, fig.width=10, fig.align="center"}
gg0 = ggplot() + geom_polygon(data = usa, aes(x=long, y = lat, group = group)) + 
  coord_fixed(1.3)
gg1 <- ggplot() + 
  geom_polygon(data = usa, aes(x=long, y = lat, group = group), fill = "violet", color = "blue") + 
  coord_fixed(1.3)
labs <- data.frame(
  long = c(-122.064873, -122.306417),
  lat = c(36.951968, 47.644855),
  names = c("SWFSC-FED", "NWFSC"),
  stringsAsFactors = FALSE
)  
gg2 = gg1 + 
  geom_point(data = labs, aes(x = long, y = lat), color = "yellow", size = 2)

grid.arrange(gg0, gg1, gg2, nrow = 1)
```

* Key points
  + "group = group" is essential, you can try code without the group argument
  + coord_fixed(): It fixes the relationship between one unit in the y direction and one unit in the x direction. 



## US state
```{r, echo=TRUE, fig.width=10, fig.align="center", warning = F}
### Draw plots for all USA States
states <- map_data("state")
ggplot(data = states) + 
  geom_polygon(aes(x = long, y = lat, fill = region, group = group), color = "white") + 
  coord_fixed(1.3) + guides(fill=FALSE) 

### Draw a subset of USA states
### Use function subset to select states. 
west_coast <- subset(states, region %in% c("california", "oregon", "washington"))
ggplot(data = west_coast) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = "palegreen", color = "black") + 
  coord_fixed(1.3)
```

Can you tell the difference between *color* and *fill* argument? 

# Reference
For more information about ggplot2, see https://r4ds.had.co.nz/data-visualisation.html.

