---
title: 'Week 3: Visualisation of Spatial Data -- Melbourne Example'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
</div>


# Goal


* In this lab, you will learn 
  + Reading a map (and data) through shapefile. The example map is about all local government areas (LGA) within  Metropolitan Melbourne Area. 
  + The shapefile will be stored as SpatialPolygonsDataFrame in R. Some manipulation of SpatialPolygonsDataFrame will be presented. 
  + The visualisation of SpatialPolygonsDataFrame.



# Shapefile

* A shapefile is a simple, nontopological format for storing the geometric location and attribute information of geographic features. It usually includes the following file
  + .shp:  	geometries
  + .shx:  	an index file to the geometries
  + .dbf:	  storing attribute data


* It is developed by Esri (Environmental Systems Research Institute, the company makes famous ArcGIS software). In R, if you want to read a shapefile, you need to use function *readOGR* in package *rgdal*.



## Melbourne Example

The shapefile is available in folder *Mel* in Canvas, and it is downloaded from https://www.igismap.com/australia-shapefile-download/ (This website also has shapefile of other Australia regions and cities). You can download all LGA in Victoria through  https://data.gov.au/dataset/ds-dga-bdf92691-c6fe-42b9-a0e2-a4cd716fa811/details (However, it does not indicate which LGA belongs to Greater Melbourne). 

The shapefile here is named *Aus-Melbourne02*, which records information from all LGA in Greater Melbourne. In readOGR, for more details, see help file using *?readOGR* 

*The package "rgdal" can be installed in my Windows and Mac OS system. For Mac, you also need to install Xcode and XQuartz. If you are using Ubuntu 16.04, Rstudio may have trouble to install this package. (I am sorry that I cannot fix this problem since I am not very familiar with Linux system. There is some informaiton on Internet, which may help. If it does not work, you need to find a Windows/Mac system, such as in G69.) *

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
library(rgdal)
library(sp)
gmel = readOGR(dsn = "datasets/Mel/Aus-Melbourne02.shp", "Aus-Melbourne02")
```

```{r, echo=TRUE, fig.width=6, fig.align="center"}
class(gmel) 
gmel = gmel[-1,]  # the first element is greater melbourne, not LGA
```

The shapefile you just read into is stored in *gmel*, which is a SpatialPolygonsDataFrame class. The olinda data includes all cities (LGA) within Metropolitan Melbourne Area.  Attributes of each neighbourhood is stored in *gmel@data*, and the geometric shape of each neighbourhood is stored as SpatialPolygons, which can be accessed using *as(gmel, "SpatialPolygons"*

## Manuplation of SpatialXXXDF

R package *mapview* provides functions to create interactive visualisation of spatial data.(For more detailed tutorial, please look at "https://r-spatial.github.io/mapview/index.html").  
A simple interactive map can generated as
```{r, echo=TRUE, fig.width=10, fig.align="center"}
library(mapview)
mapView(gmel)
```
I find out that by drawing SpatialPolygons directly on maps, it is easier to obtain a feeling of the dataset. Please click around the map, a lot of information will show up.  

Variables
```{r, echo=TRUE, fig.width=6, fig.align="center"}
gmel$name                  # Extracting a variable
gmel$ID = 1:31              # Add a new variable
gmel$ID = NULL              # Delete a variable
```

Select LGA == City of Banyule is 
```{r, echo=TRUE, fig.width=6, fig.align="center"}
i = which(gmel$name == "City of Banyule")
g = gmel[i, ]
mapView(g)
```



# Combining Data (Optional)

*If you find this part difficult, you can skip this part and move to the second part directly. It is optional, and will not be tested in exam.*


```{r, echo=TRUE, fig.width=6, fig.align="center"}
class(gmel) 
gmel = gmel[-1,]  # the first element is greater melbourne, not LGA
```


In https://discover.data.vic.gov.au/dataset/2015-local-government-area-profiles, many statistics for each LGA in Victoria is recorded on 2015, and it is available in *LGAData.csv*. It includes around 400 variables. 

* Here, we choose
  + name: LGA name
  + price: median house price
  + dis: distance to Melbourne (in km)
  + off: totoal offences per 1000 population
  + inc: median household income weekly


```{r, echo=TRUE, fig.width=6, fig.align="center"}
LGAData = read.csv("datasets/LGAData.csv")
LGAsub = subset(LGAData, select = c("LGA.Name","Median.house.price", "Distance.to.Melbourne", "Total.offences.per.1.000.population", "Median.household.income"))
colnames(LGAsub) = c("name","price", "dis", "off", "inc")
```

We want to combine the data in *LGAsub* and *gmel* together. However, the use different names. For example, in *LGAsub*, it says "City of Banyule", while it says "Banyule (C)" in *gmel*. We want to add a new variable *name2*, and for the above LGA, it will be named "Banyule".  Therefore, *LGAsub* and *gmel* will be matched and joined. 
```{r, echo=TRUE, fig.width=6, fig.align="center"}
# Add name2 for gmel
gmel$name2 = NA
for (i in 1:length(gmel$name)){
  lganame = as.character(gmel$name[i])
  gmel$name2[i] = (strsplit(lganame," of ")[[1]])[2]
}

LGAsub$price = as.numeric(gsub('[$,]', '', LGAsub$price))  # Change price from dollar form to numbers
LGAsub$inc = as.numeric(gsub('[$,]', '', LGAsub$inc))       # Change income from dollar form to numbers
LGAsub$dis = as.numeric(gsub(" km", "", LGAsub$dis))       # distance to numeric

# Add name2 for LGAsub
LGAsub$name2 = NA
for (i in 1:dim(LGAsub)[1]){
  lganame = as.character(LGAsub$name[i])
  LGAsub$name2[i] = substr(lganame, 1, nchar(lganame)-4)
}

temdata = subset(gmel@data, select = "name2") 
library(plyr) 
meldata = plyr::join(temdata, LGAsub, by = "name2", type="left")
gmel2 = gmel
gmel2@data = meldata
# Save the datasets
save(gmel2, file = "datasets/gmel2.Rdata")   
```
Here, *plyr::join* means the *join* function in package *plyr*, since another package *dplyr* also has a function called *join*. 





# Visualisation using spplot


If you have not done the previous section, you can download the data *gmel2*
```{r, echo=TRUE, fig.width=6, fig.align="center"}
load("datasets/gmel2.Rdata")
```

The visualisation can be achieved using *spplot*.
```{r, echo=TRUE, fig.width=6, fig.align="center",warning = F}
spplot(gmel2, "price", col.regions = grey.colors(20, 0.9, 0.3))
```

* Code
  + The first argument shall be a SpatialXXX, or SpatialXXXDF. In this case, it is a SpatialPolygonsDataFrame
  + The second argument is the value you want to draw. In this case, it is DEPRIV.
  + The third argument is the color scheme. Other types of colors
    + *spplot(gmel2, "price", col.regions = grey.colors(20, 0.9, 0.3), cut=4)*
    + *spplot(gmel2, "price", col.regions = heat.colors(20,alpha=0.8))*
    + *spplot(gmel2, "price", col.regions = rev(heat.colors(20,alpha=0.8)))*




\newpage


# References
* For more information, please see (Not required to read, and treat the following materials as reference. )
  + Classes and Methods for Spatial Data: the sp Package (https://cran.r-project.org/web/packages/sp/vignettes/intro_sp.pdf)
  + Introduction to visualising spatial data in R
(https://cran.r-project.org/doc/contrib/intro-spatial-rl.pdf)
  + Spatial data manipulation
(https://rspatial.org/raster/spatial/index.html)

