---
title: "Week 1: Data Transformation -- RDS -- dplyr  tidyverse"
---

<div style="margin-bottom:50px;">

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>





* Please study Chapter 3 of the book *R for Data Science (2nd edition)* (https://r4ds.hadley.nz). 

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
library(dplyr)
library(nycflights13)
library(tidyverse)
```



# tibble 
```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
flights
```


``flights is a tibble, a special type of data frame used by the tidyverse to avoid some common gotchas. The most important difference between tibbles and data frames is the way tibbles print; they are designed for large datasets, so they only show the first few rows and only the columns that fit on one screen."


# 3.1.3 

From the book: 
* it’s worth stating what they have in common:
  + 1  The first argument is always a data frame.
  + 2 The subsequent arguments typically describe which columns to operate on using the variable names (without quotes).
  + 3 The output is always a new data frame.

```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
flights |>
  filter(dest == "IAH") |> 
  group_by(year, month, day) |> 
  summarize(
    arr_delay = mean(arr_delay, na.rm = TRUE)
  )
```

# Other parts


3.2 Rows: filter, arrange, distinct, count

3.3 Columns: mutate, select, rename, relocate

3.5 Groups: group_by, summarize


# Optional

If you have time, you can read the following chapter, such as 5. Data tidying