---
title: "Week 8: Geostatistics--Kriging"
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>




This Lab corresponds to Cp 4: 9


# Introduction

* In this Lab, we will learn two key functions
  + krige.control: specify the parameter for kriging, no computing involved.
  + krige.conv: to compute 

\vspace{5pt}

* Usage
  + krige.control(type.krige = "ok", obj.model = NULL)
  + Type of krige: "sk", "ok" (default)
  + obj.model: typically an output of *likefit*

\vspace{5pt}

* Usage
  + krige.conv(geodata, locations, krige)
  + geodata
  + locations: $\mathbf s_0$, data.frame, multiple locations. 
  + krige: output of krige.control.





# Ordinary Kriging


First, we fit the model $$Z(\mathbf{s}) = \mu + U(\mathbf{s}) + \nu(\mathbf{s}).$$ Here *trend = "cte"* specifies that the trend $\mu(\mathbf{s})$ is a constant ($\mu(\mathbf{s})=\mu$). 
```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
library(geoR)
data(ca20)
fit = likfit(ca20, ini = c(100, 120), trend = "cte")
```

Secod, we setup the Ordinary Kriging. 
```{r echo=TRUE, message=FALSE, warning=FALSE, hold=FALSE}
pred_locs = expand.grid(x = seq(5300, 5800, 10), y = seq(5000,5600, 10))   # new locations s_0 created
head(pred_locs)     ## Here are the first six locations
KC = krige.control(obj.model = fit)
```

Third, we make predictions on all locations on *gr*
```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
ca20pred = krige.conv(ca20, loc = pred_locs, krige = KC)
```
* The output includes
  + ca20pred$predict: the kriging estimates for all locations specified at *loc*. 
  + ca20pred$krige.var: the kriging variance for all locations specified at *loc*.


Last, we visualise both Kriging estimates and Kriging variances. The function to use is *geom_tile* 

```{r echo=TRUE, results="hold", warning=FALSE}
ca20_res = data.frame(x = pred_locs[,1], y = pred_locs[,2], 
                      krige = ca20pred$predict, krige_var = ca20pred$krige.var)


library(ggplot2)
ggplot(ca20_res) + geom_tile(aes(x = x, y = y, fill = krige))
```

Kriging Variances

```{r echo=TRUE, results="hold", warning=FALSE}
ggplot(ca20_res) + geom_tile(aes(x = x, y = y, fill = krige_var))
```



Kriging weights

```{r echo=TRUE, results="hold", warning=FALSE}
coords = cbind(c(0.2, 0.25, 0.6, 0.7), 
               c(0.1, 0.8, 0.9, 0.3))   # data
KC = krige.control(ty = "ok", cov.model = "mat", 
                   kap = 1.5, nug = 0.1, 
                   cov.pars = c(1, 0.1))    # model parameter
# Weights
krweights(coords, c(0.5, 0.5), KC)                              
```




# Universal Kriging


<span style="color: red;">Both ordinary kriging and universal kriging</span> using setting  *type.krige = "ok"*. Once you put the trend in, the function will distinguish two types of kriging.

## Universal Kriging: Polynomial of coordinates 

* We need to provide $x_1(\mathbf s_0), \ldots, x_p(\mathbf s_0)$ for the function
  + in function *krige.control*
  + trend.d: specify data trend
  + trend.l: specify prediction trend

\vspace{10pt}

Work with polynomial. 

```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
fit = likfit(ca20, ini = c(100, 60), trend = "1st")
gr = pred_grid(ca20$borders, by = 10)
KC = krige.control(obj.model = fit, trend.l="1st", 
                   trend.d="1st")
ca20pred = krige.conv(ca20, loc = gr, krige = KC)
```


## Universal Kriging: covariates 

```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
fit2 = likfit(ca20, ini = c(100, 60), trend = ~altitude)
gr = pred_grid(ca20$borders, by = 10)
graltitude = rnorm(10908,mean=5.5)
KC2 = krige.control(type.krige = "ok", obj.model=fit2, 
                    trend.d=~altitude, trend.l=~graltitude)
ca20pred2 = krige.conv(ca20, loc = gr, krige = KC2)
```
In universal kriging, we need to know $x(\mathbf s_0)$. In the above code, I randomly generated $x(\mathbf s_0)$ from *graltitude = rnorm(10908,mean=5.5)*. In practice, you shall know $x(\mathbf s_0)$ in advance. 

# Exercise

You can also directly using Kriging formulas from slides to calculate the Kriging estimator and the Kriging variance. If you have time, try to calculate them using the simple Kriging for the location $(5300, 5000)$. 