---
title: 'Week 5: Spatial Point Process -- spatstat'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>




# Test of CSR


## Generate Poisson Process: HPP


```{r echo=TRUE, results="hold", message=FALSE}
library(spatstat)
data1 = rpoispp(lambda=100, win=owin(c(0,2),c(0,2)))
plot(data1, pch=46, cex =3)                           
```

* Argument: 
  + lambda=100: intensity function, <span style="color: red;">NOT number of points</span>
  + win=owin(c(0,2),c(0,2)): spatial point process data will be generated in the domain D = [0,2]*[0,2].
  + Question: What is the expected number of events?




## Quadrat test

Here, we will use R package *spatstat*

```{r echo=TRUE, warning=FALSE, message = FALSE}
quadrat.test(data1, nx = 4, ny= 4, alternative = "two.sided", method = "Chisq")
```
* Argument: 
  + pp: the dataset must be a "ppp" class.
  + <span style="color: red;">nx=4, ny =4</span>: number of block on x-axis and y-axis. Here, we have 4*4=16 blocks. 
  + alternative: alternative hypothesis. You can choose one of "two.sided", "regular", "clustered"
  + method: the method used to compute p-value. "Chisq" (analytical way), "MonteCarlo"
  
Since p-value is bigger than 0.05, we fail to reject $H_0$.


  
## Test: G Function

```{r echo=TRUE, warning=FALSE, message = FALSE}
r=seq(0,sqrt(2), by=0.005)
Gtest=envelope(data1,fun=Gest, r=r, nrank=2, nsim=99)
```
* Argument: 
  + nsim: the number of simulated datasets
  + nrank: similar to deciding the significant level
    + Note that (2+2+1)/(99+1) = 0.05, and therefore, it is a similar to test at <span style="color: red;">"significant level"</span> = 0.05 
    + That is, if you want to ensure $\alpha$ (significant level), you need to set $(2nrank+1)/(nsim+1) = \alpha$. 
  + fun: Gest
  + r: where the G function will be evaluated. ($h$ in our slides.)



```{r echo=TRUE, warning=FALSE, message = FALSE}
plot(Gtest)
```

What do you conclude from the above plot?



## ppp class

As mentioned earlier, package *spatstat* uses ppp class. We can transform between *SpatialPoints* class in *sp* package and *ppp* class using the function *as* in package *maptools*

```{r message=TRUE, echo=TRUE, warning=FALSE}
library(sp)
library(maptools)
# from ppp class to SpatialPoints class 
spdata1=as(data1, "SpatialPoints")
class(spdata1)  
# from SpatialPoints class to ppp class
pppdata1=as(spdata1, "ppp")
class(pppdata1) 
```

# Spatial Point Models




## Create IPP

```{r echo=TRUE, results="hold"}
# inhomogeneous Poisson process in unit square
# with intensity lambda(x,y) = min(10 * exp(3*x), 100), where $s = (x, y)^T$
lamb = function(x,y) { 10 * exp( 3 * x)}
ipp = rpoispp(lambda=lamb, lmax=100) 
plot(ipp, pch=46, cex =3)
```

Here, *lmax* specifies the largest possible intensity. For example, when $x=1$, we have  $10\exp(3x)=201$. However, since *lmax = 100* is specified, the intensity will be 100. That is, we achieve "intensity lambda(x,y) = min(10 * exp(3*x), 100)". 



## Create Neyman-Scott Process: clustered data pattern

```{r echo=TRUE, results="hold"}
nclust <- function(x0, y0, radius, n) {
          return(runifdisc(n, radius, centre=c(x0, y0)))
            }
rNS = rNeymanScott(kappa=10, expand=0.2, 
                   rcluster = nclust, radius=0.2, n=5)
plot(rNS, pch=46, cex =3)
```

* Argument: 
  + kappa: Intensity of the Poisson process for parent points (that is, $\lambda$ in our slides) -- step (i) 
  + expand: Size of the expansion of the simulation window for parent points -- step (i) 
  + rcluster: step $(ii^*)--(iii^*)$: here runifdisc function is used.

\vspace{10pt}

* About runifdisc function: runifdisc(n, radius)
  + Generate a random point pattern containing $n$ independent uniform random points in a circular disc.
  + n: Number of offspring. Here, n=5 -- in step $(ii^*)$. 
  + radius: Radius of the circle. radius = 0.2 
  + center: Coordinates of the center of the circle (that is, the location of the parent).  
  
## Create Matern I and Matern II: regular data pattern
```{r echo=TRUE, results="hold"}
matI = rMaternI(kappa = 10, r = 0.1)
matII = rMaternII(kappa = 10, r = 0.1)
par(mfrow=c(1,2))
plot(matI, main="", pch=46, cex =3)
plot(matII, main="", pch=46, cex =3)
```

* kappa: the intensity of the Poisson process (that is, $\lambda$ in our slides). 
* r: Inhibition distance. ($\delta$ in our slides)


## Further Study

If you have time, can you test *rNS* data whether it follow clustered point pattern?