---
title: "Week 11: Visulization of Spatio-Temporal Data (Optional)"
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>





# Introduction

* In Lab 11, we follow Lab 2.2 of *Spatio-Temporal Statistics with R*. You can download the electronic pdf through https://spacetimewithr.org/. 
  + In this document, I will only highlight some important points; that is, the visulization of spatio-temporal data.  


# Spatial Plots
* First, we load the dataset.
  + z: temperature (F)
  + id: weather station ID. 
  + t: the same as the column *day*
  + We do not really used column *julian* nor *proc*

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE, warning = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)


load("datasets/Tmax.Rdata")
head(Tmax)
```

* Next, the plot will be produced by package *ggplot2*. The new function here

We can also use *sf* object from Lab 5 and Lab 6 to achieve the same results. We transform the data to *sf* object using the function *st\_as\_sf*. Here, 4326 is WGS 84, which is the longitude and latitude. 

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
Tmax_1 = Tmax %>% filter(t %in% c(1, 15, 30))  # extract data

state_sf = st_as_sf(maps::map("state", plot = FALSE, fill = TRUE), 
                        crs = st_crs(4326))
Tmax_1sf = st_as_sf(Tmax_1, coords = c("lon", "lat"), crs = st_crs(4326))
```

Next, visualise the data using *geom\_sf*

```{r, echo=TRUE, fig.width=8, fig.align="center", message = FALSE}
NOAA_plot2 = ggplot() + geom_sf(data = state_sf) + 
  geom_sf(data = Tmax_1sf, aes(color = z)) + 
  coord_sf(xlim = c(-105, -75), ylim = c(25, 50)) + 
  scale_colour_gradientn(name = "degF", colours = terrain.colors(10)) + 
  facet_grid(~date)


NOAA_plot2
```





# Time-Series Plots

First, we randomly select 10 station for visualization. 

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
UIDs <- unique(Tmax$id)                     # extract IDs
UIDs_sub <- sample(UIDs, 10)                # sample 10 IDs
Tmax_sub <- filter(Tmax, id %in% UIDs_sub)  # subset data
head(Tmax_sub)
```

Next, time-series is plotted. 

```{r, echo=TRUE, fig.width=8, fig.align="center", message = FALSE}
## ------------------------------------------------------------------------
TmaxTS <- ggplot(Tmax_sub) +
    geom_line(aes(x = t, y = z)) +          # line plot of z against t
    facet_wrap(~id, ncol = 5) +             # facet by station
    xlab("Day number (days)") +             # x label
    ylab("Tmax (degF)") +                   # y label
    theme_bw() +                            # BW theme
    theme(panel.spacing = unit(1, "lines")) # facet spacing: the space between each subplots. 


TmaxTS
```


# Spatio-Temporal Processess Animation

```{r, fig.width = 10, fig.height = 6, out.width = "50%", message = FALSE}
library(stpp)
library(splancs)
```

```{r, fig.width = 8, fig.height = 3}
data <- read.table("datasets/AEGISS_ixyt.dat", header = TRUE)
data <- data[order(data$t), ]
xy <- cbind(data$x, data$y)
poly <- scan("datasets/AEGISS_poly.dat", what = character())
poly <- as.numeric(poly)
poly <- matrix(poly, 120, 2, byrow = TRUE)
```


```{r, fig.width = 8, fig.height = 3}
quartz() # open a new window and we can see the animation. Use windows() for windows system. 
xyt_data <- as.3dpoints(data$x[1:1000], 
                        data$y[1:1000], 
                        data$t[1:1000])
stpp::animation(xyt = xyt_data, 
                s.region = poly, 
                t.region = xyt_data[,3], runtime = 5)
```
