---
title: 'Week 10: Spatial Lag Model'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>



# Median House Price Example

* The dataset is collected on all LGA in great Melbourne in 2015 by Government. The variables include
  + name: LGA name
  + price: median house price
  + dis: distance to Melbourne (in km)
  + off: totoal offences per 1000 population
  + inc: median household income weekly

*The goal of this example is not to analyze Melbourne housing price. The goal is to show you an example using spatial lag model. I prefer to use (if possible) some example close to Melbourne/Victoria rather than Crime data in ohio, USA.*



```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE, warning = F}
library(sp)
library(sf)
library(ggplot2)
library(mapview)
```

Here is some data visualisation
```{r, echo=TRUE, fig.width=6, fig.align="center"}
load("datasets/gmel2.Rdata")
mapView(gmel2, zcol = "price")
```




# Spatial Lag Model

Before 2020, the function lagsarlm is in the package *spdep*. Now it is moved to *spatialreg* package, and I suggest you to use *spatialreg::lagsarlm* to make sure you are using *lagsarlm* function from *spatialreg* package

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE, warning = F}
library(spdep)
library(spatialreg)
```


```{r, echo=TRUE, fig.width=6, fig.align="center"}
gmel2.sp = as(gmel2, "Spatial")
swpoly = poly2nb(gmel2.sp)
## row sum standarize to 1
colw = nb2listw(swpoly, style = "W")

## This is a demonstration on how to use function "lagsarlm"
## Not really about how to analyze melbourne housing in depth. 
mlag = spatialreg::lagsarlm(log(price)~dis+off + inc, data=gmel2, 
                listw=colw) 

```


The output

```{r, echo=TRUE, fig.width=6, fig.align="center", hold=TRUE}
# Just like lm output. distance and income are significant, while offense is not.
summary(mlag)
```


```{r, echo=TRUE, fig.width=6, fig.align="center", results="hold"}
## beta estimates and asymptotic standard error
mlag$coefficients
mlag$rest.se
```

```{r, echo=TRUE, fig.width=6, fig.align="center", results="hold"}
## rho estimates and asymptotic standard error
mlag$rho
mlag$rho.se
```

```{r, echo=TRUE, fig.width=6, fig.align="center", results="hold"}
## sigma^2 estimates
mlag$s2

## AIC value
AIC(mlag)

## The loglikelihood value
mlag$LL
logLik(mlag)

## the covariance matrix of the estimator ()
vcov(mlag)
```




