---
title: 'Week 7: Visualisation of Spatial Data--Australian maps'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>




# Australia maps




## Package ozmaps  
* It provide the following map (all in form of sf class):
  + ozmaps::ozmap_states: state boundary
  + ozmaps::abs_ced: Commonwealth Electoral Divisions
  + ozmaps::abs_lga: Local Government Areas
```{r, echo=TRUE, fig.width=6, fig.align="center"}
library(sp)
library(sf)
library(ggplot2)

oz_states = ozmaps::ozmap_states
ggplot(oz_states) + 
  geom_sf() + 
  coord_sf()

oz_votes = ozmaps::abs_ced
ggplot() + 
  geom_sf(data = oz_votes) + 
  coord_sf()

lga = ozmaps::abs_lga
ggplot() + 
  geom_sf(data = oz_states, mapping = aes(fill = NAME), show.legend = FALSE) +
  geom_sf(data = lga, fill = NA) + 
  coord_sf() 
```
 
## Victoria maps

* Most shapefile can found in https://data.gov.au. 
+ All LGA: VICLGA folder -- 92 elements
+ All suburb: VICSub folder  -- 2973 elements
+ LGA in Greater Melbourne -- 32 element (1st element is greater melbourne)

All datasets are read into as SpatialPolygonsDataFrame
```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
library(rgdal)
library(sp)
VICLGA = readOGR(dsn = "datasets/VICLGA/VIC_LOCALITY_POLYGON_shp.shp", "VIC_LOCALITY_POLYGON_shp")
VICSub = readOGR(dsn="datasets/VICSub/VIC_LOCALITY_POLYGON_shp.shp", "VIC_LOCALITY_POLYGON_shp")
MELLGA = readOGR(dsn = "datasets/Mel/Aus-Melbourne02.shp", "Aus-Melbourne02")
```




* You will cut SpatialPolygons you wants -- Essentially manipulation of SpatialPolygons.
* One useful funciton in *crop* in package *raster*
  + crop(x, y)
+ crop returns a geographic subset of an object *x* as specified by an Extent object *y*
  + ?crop for more details

* Here we will create
+ Task 1: all surburbs in Greater Melbourne
+ Task 2: all surburbs in City of Melbourne

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
library(raster)
library(mapview)
# Before crop, first check CRS
VICLGA@proj4string
VICSub@proj4string
MELLGA@proj4string

# Making all CRS the same
crs0 = VICLGA@proj4string
MELLGA = spTransform(MELLGA, crs0)
MELLGA@proj4string

## Task 1
GMEL = MELLGA[1, ]   # Greater Melbourne
GMELSub = crop(VICSub, GMEL)
mapView(GMELSub)      

## Task 2: this code takes a while to run. 
i = which(MELLGA$name == "City of Melbourne")
CMEL = MELLGA[i, ]
#CMELSub = crop(VICSub, CMEL)      
#mapView(CMELSub) 
```



* There are more shapefiles in *PlanMelbourne* Folder, which can be downloaded from https://www.planmelbourne.vic.gov.au/maps/spatial-data
+ I have not explore all of shapefiles
+ It is quite useful. 
+ The following gives *regions of Victoria* and *CBD*
  
```{r, echo=TRUE, fig.width=6, fig.align="center"}
#SReg = readOGR(dsn="datasets/PlanMelbourne/Administrative/State Regions_region.shp", "State Regions_region")

#library(mapview)
#mapView(SReg)
#CBD = readOGR(dsn="2_ggplot/PlanMelbourne/Central City Urban Renewal and Precincts/Central Business District_region.shp", "Central Business District_region")
#mapView(CBD)
```