---
title: 'Week 11: Visulization of Spatio-Temporal Data (Optional)'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>





# Introduction

* In Lab 11, we follow Lab 2.2 of *Spatio-Temporal Statistics with R*. You can download the electronic pdf through https://spacetimewithr.org/. 
  + In this document, I will only highlight some important points; that is, the visulization of spatio-temporal data.  


# Spatial Plots
* First, we load the dataset.
  + z: temperature (F)
  + id: weather station ID. 
  + t: the same as the column *day*
  + We do not really used column *julian* nor *proc*

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE, warning = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(sf)


load("datasets/Tmax.Rdata")
head(Tmax)
```

* Next, the plot will be produced by package *ggplot2*. The new function here
  + geom_path: add US states map. You can replace these US map with Australia ones from Week 7 Lab. 
  + facet_grid(~date): for each date (1, 15 and 30), produce a plot. 
  

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
Tmax_1 = Tmax %>% filter(t %in% c(1, 15, 30))  # extract data

# Spatial Plots
NOAA_plot <- ggplot(Tmax_1) +                              # plot points
  geom_point(aes(x = lon,y = lat,                          # lon and lat
                 colour = z),                              # attribute color
             size = 1.2) +                                 # point size
  scale_colour_gradientn(name = "degF",                    # attach color scale
                        colours = terrain.colors(10)) +            
  xlab("Longitude (deg)") +                                # x-axis label
  ylab("Latitude (deg)") +                                 # y-axis label
  geom_path(data = map_data("state"),                      # add US states map
            aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  facet_grid(~date) +                                      # facet by time
  coord_fixed(xlim = c(-105, -75),
              ylim = c(25, 50))  +                         # zoom in
  theme_bw()  

NOAA_plot
```

We can also use *sf* object from Lab 5 and Lab 6 to achieve the same results

First, transform the data to *sf* object using the function *st\_as\_sf*. Here, 4326 is WGS 84, which is the longitude and latitude. 

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
state_sf = st_as_sf(maps::map("state", plot = FALSE, fill = TRUE), 
                        crs = st_crs(4326))
Tmax_1sf = st_as_sf(Tmax_1, coords = c("lon", "lat"), crs = st_crs(4326))
```

Next, visualise the data using *geom\_sf*
```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
ggplot() + geom_sf(data = state_sf) + 
  geom_sf(data = Tmax_1sf, aes(color = z)) + 
  coord_sf(xlim = c(-105, -75), ylim = c(25, 50)) + 
  scale_colour_gradientn(name = "degF",                    # attach color scale
                         colours = terrain.colors(10)) + 
  facet_grid(~date)
```

# Time-Series Plots

First, we randomly select 10 station for visualization. 

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
UIDs <- unique(Tmax$id)                     # extract IDs
UIDs_sub <- sample(UIDs, 10)                # sample 10 IDs
Tmax_sub <- filter(Tmax, id %in% UIDs_sub)  # subset data
head(Tmax_sub)
```

Next, time-series is plotted. 

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
## ------------------------------------------------------------------------
TmaxTS <- ggplot(Tmax_sub) +
    geom_line(aes(x = t, y = z)) +          # line plot of z against t
    facet_wrap(~id, ncol = 5) +             # facet by station
    xlab("Day number (days)") +             # x label
    ylab("Tmax (degF)") +                   # y label
    theme_bw() +                            # BW theme
    theme(panel.spacing = unit(1, "lines")) # facet spacing: the space between each subplots. 


TmaxTS
```


