---
title: 'Week 6: Visualisation of Spatial Data (3) -- Maps'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>



# Victoria maps: geom_sf


All datasets are read into as *sf* objects
```{r, echo=TRUE, fig.width=6, fig.align="center", message = F}
library(ggplot2)
library(sf)
VICLGA = st_read(dsn = "datasets/VICLGA/VIC_LOCALITY_POLYGON_shp.shp", "VIC_LOCALITY_POLYGON_shp")
VICSub = st_read(dsn="datasets/VICSub/VIC_LOCALITY_POLYGON_shp.shp", "VIC_LOCALITY_POLYGON_shp")
```



First, we make sure that all *sf* objects has the same CRS
```{r, echo=TRUE, fig.width=6, fig.align="center", message = F}
## Retrieve CRS
st_crs(VICLGA) 
st_crs(VICSub)

## Transform crs system to WGS84(epsg: 4326)
VICLGA <- st_transform(VICLGA, 4326)
VICSub <- st_transform(VICSub, 4326)

## Check CRS
#st_crs(VICLGA) 
#st_crs(VICSub)
```

Second, we want to obtain all suburbs in City of Banyule. Note that we already have all suburbs in Victoria, and we also know the boundary of City of Banyule from VICLGA. We only need to ``cut" all suburbs from *VICSub* using boundary of City of Banyule. 

* One useful funciton in *st_intersection* in package *sf*
  + st_intersection(x, y)
  + this function returns a geographic subset of an object *x* as specified by an extent object *y*

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE, warning = F}
library(raster)
library(mapview)


## Obtain all suburbs in City of Banyule (including part of suburb)
i = which(VICLGA$VIC_LGA__3 == "BANYULE")
CMEL = VICLGA[i, ]
tmp = st_intersection(VICSub, CMEL)
tmp$VIC_LOCA_2
```


The object *tmp* includes suburbs which is not in City of Banyule, such as Balwyn North. Here is a list of all suburbs in City of Banyule (https://en.wikipedia.org/wiki/City_of_Banyule). The reason for this is likely due to the different resolution and accuracy of shapefiles (VICSub and VICLGA).   

Therefore, I will remove all the suburbs with a very small area (Here, the unit is square meter, and 10000 is small by suburb standards.)  

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
## Obtain all suburbs in City of Banyule (including part of suburb)
tmp$area = st_area(tmp)
BSub = tmp[as.numeric(tmp$area) > 100000,]
mapView(BSub)
save(BSub, file = "datasets/BSub.Rdata")   
```



There are major activity centres. According to the goverment, *These are places that provide a suburban focal point for services, employment, housing, public transport and social interaction. They have different attributes and provide different functions, with some serving larger subregional catchments. Plan Melbourne identifies 121 existing and future Major Activity Centres across Melbourne.*

Suppose we want to obtain all major activity centres in the City of Banyule. Note these acitivity centres are represented as *points*

```{r, echo=TRUE, fig.width=7, fig.align="center", message = FALSE}
centres = st_read(dsn = "datasets/PlanMel/Activity Centres/ActivityCentrePoints_font_point.shp", 
                 "ActivityCentrePoints_font_point")
centres <- st_transform(centres, 4326)
## obtain all centres within the City of Banyule (CMEL)
tmp = st_within(centres, CMEL, sparse = F)
Bcen = centres[tmp[,1],]

library(ggplot2)
library(sf)
ggplot() + geom_sf(data = BSub, aes(fill = VIC_LOCA_2)) + 
  geom_sf_label(data = BSub, aes(label = VIC_LOCA_2), size = 2.1) + 
  geom_sf(data = Bcen, size = 4) + theme(legend.position="none")
```




Suppose you also want to draw major roads in City of Banyule. The shapefile of all roads are in the folder *VMTRANS*. Here, we want to draw all roads with Road Class = 0, 1, 2. (I do not know what Road Class really means here, but it appears the smaller number indicates a more important road.)


```{r, echo=TRUE, fig.width=7, fig.align="center", message = FALSE}
road = st_read(dsn = "datasets/VMTRANS/TR_ROAD.shp", "TR_ROAD")
road <- st_transform(road, 4326)
## You can see all roads in Banyule
#mapView(road)
## We only draw road with CLASS_CODE = 0, 1, 2
highway = road[road$CLASS_CODE < 3,]
#mapView(highway)

## Now add all together
ggplot() + geom_sf(data = BSub, color = "gray") + 
  geom_sf(data = Bcen, size = 4) + 
  geom_sf(data = highway, aes(color = as.factor(CLASS_CODE)), lwd = 1.5)  
```

# R package *maps* and *mapdata*

R package *maps* and *mapdata* provide some maps. They are not in *sf* format. To convert map data to sf, use following codes
```{r, echo=TRUE, fig.width=7, fig.align="center", message = FALSE}
state_sf = sf::st_as_sf(maps::map("state", plot = FALSE, fill = TRUE))

ggplot() + geom_sf(data = state_sf)
```

If you want to plot directly using *geom_polygon*, see Extra3. 

