---
title: 'Week 10: Spatial Lag Model'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>

<span style="color: red;">Due to the Easter Break, this Lab will be held in Week 11. </span>



# Median House Price Example

* The dataset is collected on all LGA in great Melbourne in 2015 by Government. The variables include
  + name: LGA name
  + price: median house price
  + dis: distance to Melbourne (in km)
  + off: totoal offences per 1000 population
  + inc: median household income weekly


```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE, warning = F}
library(sp)
library(sf)
library(ggplot2)
library(mapview)
```

Here is some data visualisation
```{r, echo=TRUE, fig.width=6, fig.align="center"}
load("datasets/gmel2.Rdata")
mapView(gmel2, zcol = "price")
```

# Specify W
```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE, warning = F}
library(spdep)
library(spatialreg)
```

```{r, echo=TRUE, fig.width=6, fig.align="center"}
gmel2.sp = as(gmel2, "Spatial")
# argument: queen = TRUE by default
swpoly = poly2nb(gmel2.sp)
# Style: whether to standardise  
## "B": basic binary coding
## "W": row standardised (sums over all links to n)
colb = nb2listw(swpoly, style = "B")
colw = nb2listw(swpoly, style = "W")
apply(listw2mat(colb), 1, sum)
apply(listw2mat(colw), 1, sum)


# Construct W via distance
centre_tmp = st_centroid(gmel2$geometry)
centre = st_coordinates(centre_tmp)
n = dim(gmel2)[1]
dist_mat = as.matrix(dist(centre, diag = TRUE, upper = TRUE))
matW = 1/dist_mat
W1 = mat2listw(matW)
```


# Spatial Lag Model

Before 2020, the function lagsarlm is in the package *spdep*. Now it is moved to *spatialreg* package, and I suggest you to use *spatialreg::lagsarlm* to make sure you are using *lagsarlm* function from *spatialreg* package




```{r, echo=TRUE, fig.width=6, fig.align="center"}
## This is a demonstration on how to use function "lagsarlm"
## Not really about how to analyze melbourne housing in depth. 
mlag = spatialreg::lagsarlm(log(price)~dis+off + inc, data=gmel2, 
                listw=colw) 

```


* The output
  + LR test: Likelihood Ratio Test in Slide 20 of Cp6.
  + Wald: Wald Test in Slide 20 of Cp6.

```{r, echo=TRUE, fig.width=6, fig.align="center", hold=TRUE}
# Just like lm output. distance and income are significant, while offense is not.
summary(mlag)
```


```{r, echo=TRUE, fig.width=6, fig.align="center", results="hold"}
## beta estimates and asymptotic standard error
mlag$coefficients
mlag$rest.se
```

```{r, echo=TRUE, fig.width=6, fig.align="center", results="hold"}
## rho estimates and asymptotic standard error
mlag$rho
mlag$rho.se
```

```{r, echo=TRUE, fig.width=6, fig.align="center", results="hold"}
## sigma^2 estimates
mlag$s2

## AIC value
AIC(mlag)

## The loglikelihood value
mlag$LL
logLik(mlag)

## the covariance matrix of the estimator ()
vcov(mlag)
```




