---
title: 'Week 9: Geostatistics--Kriging'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>




# Introduction

* In this Lab, we will learn two key functions
  + krige.control: specify the parameter for kriging, no computing involved.
  + krige.conv: to compute 

\vspace{5pt}

* Usage
  + krige.control(type.krige = "ok", obj.model = NULL)
  + Type of krige: "sk", "ok" (default)
  + obj.model: typically an output of *likefit*

\vspace{5pt}

* Usage
  + krige.conv(geodata, locations, krige)
  + geodata
  + locations: $\mathbf s_0$, data.frame, multiple locations. 
  + krige: output of krige.control.





# Ordinary Kriging

First, we fit the model $$Z(\mathbf{s}) = \mu + U(\mathbf{s}) + \nu(\mathbf{s}).$$ Here *trend = "cte"* specifies that the trend $\mu(\mathbf{s})$ is a constant ($\mu(\mathbf{s})=\mu$). 
```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
library(geoR)
data(ca20)
fit = likfit(ca20, ini = c(100, 60), trend = "cte")
```

Secod, we setup the Ordinary Kriging. 
```{r echo=TRUE, message=FALSE, warning=FALSE, hold=FALSE}
gr = pred_grid(ca20$borders, by = 10)   # new locations s_0 created
head(gr)     ## Here are the first six locations, where the first locations is (4920, 4800). 
dim(gr)      ## There are a total of 10908 new locations to predict. 
KC = krige.control(obj.model = fit)
```

Third, we make predictions on all locations on *gr*
```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
ca20pred = krige.conv(ca20, loc = gr, krige = KC)
```
* The output includes
  + ca20pred$predict: the kriging estimates for all locations specified at *loc*. 
  + ca20pred$krige.var: the kriging variance for all locations specified at *loc*.


Last, we visualise both Kriging estimates and Kriging variances. 

```{r echo=TRUE, results="hold", warning=FALSE}
color = function(x)rev(heat.colors(x)) 
image(ca20pred, loc = gr, x.leg = c(4930, 5350)
      , y.leg = c(4790, 4840), col=color(15))
```

Kriging Variances

```{r echo=TRUE, results="hold", warning=FALSE}
image(ca20pred, loc = gr, val=sqrt(ca20pred$krige.var)
      , x.leg = c(4930, 5350), y.leg = c(4790, 4840)
      , col=color(15))
```



Kriging weights

```{r echo=TRUE, results="hold", warning=FALSE}
coords = cbind(c(0.2, 0.25, 0.6, 0.7), 
               c(0.1, 0.8, 0.9, 0.3))   # data
KC = krige.control(ty = "ok", cov.model = "mat", 
                   kap = 1.5, nug = 0.1, 
                   cov.pars = c(1, 0.1))    # model parameter
# Weights
krweights(coords, c(0.5, 0.5), KC)                              
```


Convert to SpatialPointsDataFrame


```{r echo=TRUE, results="hold", warning=FALSE}
library(sp)
gr0 = polygrid(gr, borders = ca20$border, bound = T)
capred = data.frame(gr0, ypred=ca20pred$predict)
coordinates(capred) = ~ Var1 + Var2
class(capred)                            
```
We can visualize using all functions related to SpatialPoints. 


# Universal Kriging


<span style="color: red;">Both ordinary kriging and universal kriging</span> using setting  *type.krige = "ok"*. Once you put the trend in, the function will distinguish two types of kriging.

## Universal Kriging: Polynomial of coordinates 

* We need to provide $x_1(\mathbf s_0), \ldots, x_p(\mathbf s_0)$ for the function
  + in function *krige.control*
  + trend.d: specify data trend
  + trend.l: specify prediction trend

\vspace{10pt}

Work with polynomial. 

```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
fit = likfit(ca20, ini = c(100, 60), trend = "1st")
gr = pred_grid(ca20$borders, by = 10)
KC = krige.control(obj.model = fit, trend.l="1st", 
                   trend.d="1st")
ca20pred = krige.conv(ca20, loc = gr, krige = KC)
```


## Universal Kriging: covariates 

```{r echo=TRUE, results='hide', message=FALSE, warning=FALSE, hold=TRUE}
fit2 = likfit(ca20, ini = c(100, 60), trend = ~altitude)
gr = pred_grid(ca20$borders, by = 10)
graltitude = rnorm(10908,mean=5.5)
KC2 = krige.control(type.krige = "ok", obj.model=fit2, 
                    trend.d=~altitude, trend.l=~graltitude)
ca20pred2 = krige.conv(ca20, loc = gr, krige = KC2)
```
In universal kriging, we need to know $x(\mathbf s_0)$. In the above code, I randomly generated $x(\mathbf s_0)$ from *graltitude = rnorm(10908,mean=5.5)*. In practice, you shall know $x(\mathbf s_0)$ in advance. 