---
title: 'Week 4: Moran''s I, LISA'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>

# Goal

* In this lab, you will learn 
  + Creating weight matrix
  + Conducting Moran's I and LISA


# Create Weight Matrix

The dataset *gmel2* is from Week 3 Lab, and it can be downloaded at https://github.com/TedChu/90122Lab/tree/master/datasets.
```{r, echo=TRUE, fig.width=10, fig.align="center", message = FALSE}
library(rgdal)
library(spdep)
load("datasets/gmel2.Rdata")
```



R function *poly2nb* produces a spatial weight matrix from a SpatialPolygons(DataFrame) object. 
```{r, echo=TRUE, fig.width=3, fig.height =3, fig.align="center"}
Wqueen = poly2nb(gmel2)
Wrook = poly2nb(gmel2, queen=FALSE)
plot(gmel2,border="grey60", axes=TRUE)
plot(Wqueen, coordinates(gmel2), pch=19, cex=0.6, add=TRUE,col="red")
plot(Wrook, coordinates(gmel2), pch=19, cex=0.6, add=TRUE)
```

* There are three potential class for spatial weight matrix: *nb* (neighborhood) class, *mat* (matrix) class, and *listw* class.
  + They all store the same information, but in different formats.
  + Different functions may require different formats. For example, *moran.test* requires the  *listw* class. You can find these information in help files of these functions.
  + Functions to transform between these three formats: *nb2listw*, *nb2mat*, *mat2nb*, etc. 


Can you see the difference between *queen* rule and *rook* rule?


# Moran's I test
A Moran's I test is performed on the variable *price*. More information about Moran's I codes can be found in Cp1C -- 8. 


```{r, echo=TRUE, fig.align="center"}
## plot
moran.test(x=gmel2$price, nb2listw(Wqueen), alternative="two.sided")
```

Since the p-value < 0.05, we conclude that there is spatial autocorrelation for the house price in Greater Melbourne.  


* Comments
  + The function requires *listw* form for spatial weight matrix
  + Argument *alternative*
    + What will you do if the alternative test is "There is positive spatial autocorrelation".
  + The above test is *Moran I test under randomisation*. 
    + What does this mean?
    + If you want to take analytical approach mentioned in lectures, what should you do?
    
    
# LISA

LISA is performed on *price* variable. 
```{r, echo=TRUE, fig.width=3, fig.height =3, fig.align="center"}
resI = localmoran(gmel2$price, nb2listw(Wqueen), alternative="two.sided")
round(head(resI),3)
```
* For the lisa output
  + First column: the value of local moran's I
  + Second column: the expected value (which is $-1/(n-1)$)
  + Third column: the variance of local moran's I
  + Fourth column $=\frac{(Ii - E.Ii)}{\sqrt{Var.Ii}}$
  + Fifth column: the p-value. 

&nbsp;

Next, some visualisation for the above results. 
```{r, echo=TRUE, fig.width=3, fig.height =3, fig.align="center",warning = F}
gmellisa = gmel2
gmellisa@data = data.frame(gmellisa@data, lisa=resI[,1], pvalue=resI[,5])
### Spatial plots with LISA value
### rev function is used to assign darker color to higher values. 
spplot(gmellisa, "lisa", col.regions = rev(heat.colors(16,alpha=0.8)),main="LISA value")
### Spatial plots with p values
spplot(gmellisa, "pvalue", col.regions = heat.colors(16,alpha=0.8),main="p value")

### Wants to see threshold 0.05, 0.10 for p vlaues
library(RColorBrewer)     # for function brew.pal
brks=c(0.00,0.05,0.10,0.5,0.8,1.00)
pal=brewer.pal(5, "OrRd") # we select 5 colors from the palette
spplot(gmellisa, "pvalue", col.regions=pal, at = brks,  main = "p value")
```

Suppose we want to know which LGAs have significant local spatial autocorrelation ($\alpha = 0.05$). 

```{r, echo=TRUE, fig.width=3, fig.height =3, fig.align="center",warning = F}
gmellisa@data[gmellisa$pvalue < 0.05, c(1,3,7,8)]
```


They all have positive LISA values. If you want to formally test whether there are positive local autocorrelation, you will need to change *alternative* to *greater*.



# Practice: Columbus Data
If you have time, you can analyze the following Columbus data


The dataset is about 49 neighbourhoods in Columbus, OH in Year 1980, more informaiton can be found in R by *?columbus*.

It is essential a shapefile, and the difference from olinda data is columbus data is already stored in R. Thus, to read the colubus data, you use *system.file* instead. 

After read the shapefile, you will get a SpatialPolygonsDataFrame object *columbus*. 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
columbus <- readOGR(system.file("shapes/columbus.shp", package="spData")[1])
plot(columbus, border="black",main="Neigbhoods of Columbus City")
spplot(columbus, "CRIME", col.regions = rev(heat.colors(20,alpha=0.8)))
```


