---
title: 'Week 4: Moran''s I, LISA'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>

# Goal

* In this lab, you will learn 
  + Creating weight matrix
  + Conducting Moran's I and LISA


# Create Weight Matrix

We will load the dataset *gmel2* from Week 3 Lab. 
```{r, echo=TRUE, fig.width=10, fig.align="center", message = FALSE}
library(rgdal)
library(spdep)
load("datasets/gmel2.Rdata")
```



R function *poly2nb* produces a spatial weight matrix from a SpatialPolygons(DataFrame) object. 
```{r, echo=TRUE, fig.width=3, fig.height =3, fig.align="center"}
Wqueen = poly2nb(gmel2)
Wrook = poly2nb(gmel2, queen=FALSE)
plot(gmel2,border="grey60", axes=TRUE)
plot(Wqueen, coordinates(gmel2), pch=19, cex=0.6, add=TRUE,col="red")
plot(Wrook, coordinates(gmel2), pch=19, cex=0.6, add=TRUE)
```

Can you see the difference between *queen* rule and *rook* rule?


# Moran's I test
A Moran's I test is performed on the variable *price*. 


```{r, echo=TRUE, fig.align="center"}
## plot
moran.test(x=gmel2$price, nb2listw(Wqueen), alternative="two.sided")
```

* Comments
  + The function requires *listw* form for spatial weight matrix
  + Argument *alternative*
  + The above test is *Moran I test under randomisation*. 
    + What does this mean?
    + If you want to take analytical approach mentioned in lectures, what should you do?
    
    
# LISA

LISA is performed on *price* variable. 
```{r, echo=TRUE, fig.width=3, fig.height =3, fig.align="center"}
resI = localmoran(gmel2$price, nb2listw(Wqueen), alternative="two.sided")
round(head(resI),3)
```

Visualisation
```{r, echo=TRUE, fig.width=3, fig.height =3, fig.align="center",warning = F}
gmellisa = gmel2
gmellisa@data = data.frame(gmellisa@data, lisa=resI[,1], pvalue=resI[,5])
### Spatial plots with LISA value
### rev function is used to assign darker color to higher values. 
spplot(gmellisa, "lisa", col.regions = rev(heat.colors(16,alpha=0.8)),main="LISA value")
### Spatial plots with p values
spplot(gmellisa, "pvalue", col.regions = heat.colors(16,alpha=0.8),main="p value")

### Wants to see threshold 0.05, 0.10 for p vlaues
library(RColorBrewer)     # for function brew.pal
brks=c(0.00,0.05,0.10,0.5,0.8,1.00)
pal=brewer.pal(5, "OrRd") # we select 5 colors from the palette
spplot(gmellisa, "pvalue", col.regions=pal, at = brks,  main = "p value")
```


# Practice: Columbus Data
If you have time, you can analyze the following Columbus data


The dataset is about 49 neighbourhoods in Columbus, OH in Year 1980, more informaiton can be found in R by *?columbus*.

It is essential a shapefile, and the difference from olinda data is columbus data is already stored in R. Thus, to read the colubus data, you use *system.file* instead. 

After read the shapefile, you will get a SpatialPolygonsDataFrame object *columbus*. 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
columbus <- readOGR(system.file("shapes/columbus.shp", package="spData")[1])
plot(columbus, border="black",main="Neigbhoods of Columbus City")
spplot(columbus, "CRIME", col.regions = rev(heat.colors(20,alpha=0.8)))
```


