---
title: 'Week 2: CBD Pedestrian Example for ARIMA Models'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>


This Lab corresponds to Cp 2.


# Dataset

```{r, echo=TRUE, fig.width=8, fig.align="center", message = FALSE, warning = F}
library(tidyverse) # read_csv
library(fpp3)
library(ggplot2)
library(dplyr)
library(tsibble)
library(fable)
```

* Here, index 
  + 1:2208: training
  + 2209:2376: testing


```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
load("datasets/melb_sel.Rdata")
head(melb_sel)
## For the whole dataset
autoplot(melb_sel, counts) + 
  labs(x = 'Date 2019', y = 'Counts', title = 'Hourly number of pedestrians in Melbourne Central')
## For the first two weeks
autoplot(melb_sel[1:336, ], counts) + 
  labs(x = 'Date 2019', y = 'Counts', title = 'Hourly number of pedestrians in Melbourne Central')
```

# ARIMA

First, we test whether the dataset is stationary. As you can see, KPSS fails to reject null hypothesis, meaning the dataset is stationary. 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
melb_sel[1:2208, ] %>% features(counts, unitroot_kpss)
melb_sel[1:2208, ] %>% features(counts, unitroot_ndiffs)
melb_sel[1:2208, ] %>% gg_tsdisplay(plot_type = "partial", y = counts)
```
``The p-value is reported as 0.01 if it is less than 0.01, and as 0.1 if it is greater than 0.1."



Although the *counts* is stationary, I will still consider a weekly difference + a lag 1 difference.

```{r, echo=TRUE, fig.width=8, fig.align="center"}
# To determine the seasonal differencing. 
melb_sel[1:2208, ] %>% features(counts, unitroot_nsdiffs,.period = 168)  # m = 168 for a week. Try a random number and see what happens
melb_sel$diff_week <- difference(melb_sel$counts, lag=168) %>% difference(lag = 1)
melb_sel[170:2208, ] %>% features(diff_week, unitroot_kpss)
melb_sel[170:2208, ] %>% 
  gg_tsdisplay(diff_week, plot_type='partial', lag = 48)
```

It is hard to see which ARIMA model shall be used for the *counts*, while for the *diff_week*, a possible candidates is MA(3).

* I will build three model.
  + Model 1: automatic procedure for the *diff_week*
  + Model 2: MA(3) for the *diff_week*
  + Model 3: automatic procedure for the *counts*


## Model 1 

```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
fit1 <- melb_sel[170:2208, ] %>%
  model(ARIMA(diff_week ~ PDQ(0,0,0), stepwise = TRUE))
report(fit1)
```

ACF is over the threshold limits at lag = 24.
```{r, echo=TRUE, fig.width=8, fig.align="center", results = 'hold'}
fit1 %>% gg_tsresiduals()
augment(fit1) %>% features(.innov, ljung_box, lag = 10, dof = 3)
```

## Model 2
```{r, echo=TRUE, fig.width=8, fig.align="center"}
fit2 <- melb_sel[170:2208, ] %>% 
  model(ARIMA(diff_week ~ pdq(0,0,3) + PDQ(0,0,0)))
report(fit2)
```

ACF is over the thereshold at multiple lags
```{r, echo=TRUE, fig.width=8, fig.align="center"}
fit2 %>% gg_tsresiduals()
augment(fit2) %>% features(.innov, ljung_box, lag = 10, dof = 3)
```

## Model 3

```{r, echo=TRUE, fig.width=8, fig.align="center"}
fit3 <- melb_sel[1:2208,] %>% 
  model(ARIMA(counts ~ PDQ(0,0,0), stepwise = TRUE))
report(fit3)
```

ACF is over the thereshold at multiple lags
```{r, echo=TRUE, fig.width=8, fig.align="center"}
fit3 %>% gg_tsresiduals()
augment(fit3) %>% features(.innov, ljung_box, lag = 10, dof = 3)
```
* Model 1 appears to be the best, although there are still some problems. Here is my personal opinion
  + In the real-world dataset, the result is often not as good as the example in the book (e.g. Central African Republic exports in Cp 9.7)
  + lag = 24 means some daily seasonality: maybe fixable with another difference with (lag = 24). 


# Seasonal ARIMA

* I will consider the following model
  + Model 4: (2,1,1)(0,1,0)$_{168}$ for the *counts*, it is the same model as Model 1
  + Model 5: (0,1,3)(0,1,0)$_{168}$ for the *counts*, it is the same model as Model 2
  + Model 6: automatic selection for ( ,1, )(,1, )$_{168}$ for the *counts*

## Model 4 
$\phi$'s and $\theta$s are the same as Model 1, but AIC, BIC and AICc are different. 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
fit4 <- melb_sel[1:2208, ] %>% 
  model(ARIMA(counts ~ pdq(2,1,1) + PDQ(0,1,0, period = 168)))
report(fit4)
fit4 %>% gg_tsresiduals()
augment(fit4) %>% features(.innov, ljung_box, lag = 10, dof = 3)
```  

## Model 5 
```{r, echo=TRUE, fig.width=8, fig.align="center"}
fit5 <- melb_sel[1:2208, ] %>% 
  model(ARIMA(counts ~ pdq(0,1,3) + PDQ(0,1,0, period = 168)))
report(fit5)
fit5 %>% gg_tsresiduals()
augment(fit5) %>% features(.innov, ljung_box, lag = 10, dof = 3)
```  


## Model 6 
```{r, echo=TRUE, fig.align="center"}
## The following codes take quite a while to run
#fit6 <- melb_sel[1:2208, ] %>% 
#  model(ARIMA(counts ~ pdq(,1,) + PDQ(,1,, period = 168)))
#report(fit6)
#fit6 %>% gg_tsresiduals()
#augment(fit6) %>% features(.innov, ljung_box, lag = 10, dof = 3)
```  



# Forecasting

There is a function in the package to do the forecasting
```{r, echo=TRUE, fig.width=8, fig.align="center"}
forecast(fit3, h = 24) %>%
  autoplot(melb_sel[2209:2233,])
```  

I will not recommend you to use Model 1/Model 2 for the forecasting the *counts*, since it is difficult to transform the prediction interval. Instead, using the corresponding seasonal-ARIMA model if you can. 

MSE
```{r, echo=TRUE, fig.width=8, fig.align="center"}
H = 24*7
idx = 2209:(2208 + H)

tmp3 = forecast(fit3, h = H)
tmp4 = forecast(fit4, h = H)
tmp5 = forecast(fit5, h = H)
#tmp6 = forecast(fit6, h = H)
MSE3 = mean((tmp3$.mean - melb_sel$counts[idx])^2)
MSE4 = mean((tmp4$.mean - melb_sel$counts[idx])^2)
MSE5 = mean((tmp5$.mean - melb_sel$counts[idx])^2)
#MSE6 = mean((tmp6$.mean - melb_sel$counts[idx])^2)
c(MSE3, MSE4, MSE5)
```

The plots
```{r, echo=TRUE, fig.width=8, fig.align="center"}
melb_sel$tf3 = NA
melb_sel$tf4 = NA
melb_sel$tf5 = NA

melb_sel$tf3[idx] = tmp3$.mean
melb_sel$tf4[idx] = tmp4$.mean
melb_sel$tf5[idx] = tmp5$.mean

ggplot(data = melb_sel[idx, ], aes(x = date, y = counts)) + geom_line() +
  geom_line(aes(x = date, y = tf4, color = "Model 1/4: diff + auto")) + 
  geom_line(aes(x = date, y = tf5, color = "Model 2/5: diff")) + 
  geom_line(aes(x = date, y = tf3, color = "Model 3: no diff")) + 
  scale_color_manual(name = "colour", values = c("Model 1/4: diff + auto" = "red", "Model 2/5: diff" = "green", "Model 3: no diff" = "blue"))
```


Also for H = 9
```{r, echo=TRUE, fig.width=8, fig.align="center"}
H = 9
idx = 2209:(2208 + H)

tmp3 = forecast(fit3, h = H)
tmp4 = forecast(fit4, h = H)
tmp5 = forecast(fit5, h = H)
#tmp6 = forecast(fit6, h = H)
MSE3 = mean((tmp3$.mean - melb_sel$counts[idx])^2)
MSE4 = mean((tmp4$.mean - melb_sel$counts[idx])^2)
MSE5 = mean((tmp5$.mean - melb_sel$counts[idx])^2)
#MSE6 = mean((tmp6$.mean - melb_sel$counts[idx])^2)
c(MSE3, MSE4, MSE5)
melb_sel$counts[idx]
tmp4$.mean
```



# Automatic Procedure
```{r, echo=TRUE, fig.align="center"}
fitX <- melb_sel[1:2208, ] %>% 
  model(ARIMA(counts ~ pdq(,1,) + PDQ(0,1,0, period = 168)))
report(fitX)
fitY <- melb_sel[1:2208, ] %>% 
  model(ARIMA(counts ~ pdq(,0,) + PDQ(0,1,0, period = 168)))
report(fitY)
```


