---
title: 'Week 10: Spatial Lag Model'
---

<div style="margin-bottom:50px;">
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

</div>




# Median House Price Example

* The data sets is collected on all LGA in great Melbourne on 2015 by Government. Here, I will show you
  + name: LGA name
  + price: median house price
  + dis: distance to Melbourne (in km)
  + off: totoal offences per 1000 population
  + inc: median household income weekly

*The goal of this example is not to analyze Melbourne housing price. The goal is to show you an example using spatial lag model. I prefer to use (if possible) some example close to Melbourne/Victoria rather than Crime data in ohio, USA.*

*All the code (and more) will be provided in Lab 9.*


```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
library(sp)
library(sf)
library(ggplot2)
```

Here is some data the data
```{r, echo=TRUE, fig.width=6, fig.align="center"}
load("datasets/gmel2.Rdata")
head(gmel2@data)
```


```{r, echo=TRUE, fig.width=6, fig.align="center"}
class(gmel2)                 # SpatialPolygonsDataFrame
gmel2.sf = st_as_sf(gmel2)   # to sf class
library(mapview)
mapView(gmel2, zcol = "price")

```




# Spatial Lag Model

Before 2020, function lagsarlm is in *spdep*. Now it is moved to *spatialreg* package, and I suggest you to use *spatialreg::lagsarlm* to make sure you are using *lagsarlm* function from *spatialreg* package

```{r, echo=TRUE, fig.width=6, fig.align="center", message = FALSE}
library(spdep)
library(spatialreg)
```


```{r, echo=TRUE, fig.width=6, fig.align="center"}
swpoly = poly2nb(gmel2)
## row sum standarize to 1
colw = nb2listw(swpoly, style = "W")

## This is a demonstration on how to use function "lagsarlm"
## Not really about how to analyze melbourne housing in depth. 
mlag = spatialreg::lagsarlm(log(price)~dis+off + inc, data=gmel2@data, 
                listw=colw) 

```


The output

```{r, echo=TRUE, fig.width=6, fig.align="center"}
summary(mlag)
# Just like lm output. distance and income are significant, while 


## beta estiamtes and asymptotic standard error
mlag$coefficients
mlag$rest.se

## rho estimates and asymptotic standard error
mlag$rho
mlag$rho.se

## sigma^2 estimates
mlag$s2

```




